---
title: "Ran"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 0: Load necessary libraries and datasets
Load libraries that we use and the two healthcare datasets specified for the project.
We spent some time on the website reading about what's included in the datasets.

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(modelr)
library(broom)
library(readxl)
library(readr)

teen_birth <- read.csv("../Data/teen_birth.csv")

unemployment <- read_excel("../Data/Unemployment.xls", skip = 7)

education <- read_csv("../Data/Education.csv")
```
# Step 1: Tidy data into more accessible form
Using the tidyverse functions that we've learned, clean the two datasets so that data analysis wolud be more efficient.

```{r}
teen_birth2015 <- teen_birth %>%
  filter(Year == 2015) %>%
  rename(FIPS = Combined.FIPS.Code, BirthRate = Birth.Rate) %>%
  select(State, County, FIPS, BirthRate)%>%
  mutate(FIPS = as.character(FIPS))

unemployment2015 <- unemployment %>%
  rename(FIPS = FIPStxt) %>%
  select(FIPS, State, Area_name, 39:42) %>%
  rename(LaborForce = Civilian_labor_force_2015, 
         Employed = Employed_2015, 
         Unemployed = Unemployed_2015, 
         UnemploymentRate = Unemployment_rate_2015)

education20132017 <- education %>%
  select(1, 44, 45, 46, 47) %>%
  rename(Less_than_high_school = `Percent of adults with less than a high school diploma, 2013-17`,
         high_school=`Percent of adults with a high school diploma only, 2013-17`,
         some_college=`Percent of adults completing some college or associate's degree, 2013-17`,
         college=`Percent of adults with a bachelor's degree or higher, 2013-17`)
```

# Step 2: Join three datasets together
```{r}
data = teen_birth2015%>%
   inner_join(education20132017, by = c("FIPS" = "FIPS Code"))

data = data%>%
  inner_join(unemployment2015, by = "FIPS")
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
