---
title: "Project_Final"
output:
  pdf_document: default
  html_document: default
---

Group Members: Dara Zirlin, Ran Ji, Hanna Kim

We looked at the purpose codes. We found those under which the most aid was given. We then looked at the percent of aid given under these codes by year.



Below we load packages.
```{r, cache=TRUE}
#Load all necessary packages
library(tidyverse)
library(ggplot2)
library(dplyr)
```

Below we load the data.
```{r, cache=TRUE}
aiddata_purpose <- read_csv("1_Data/aiddata_purpose.csv", col_names=c("donor","recipient","year","purp_code","purp_name","usd"),skip=1)

```



Below is how much aid is given per year.
```{r}
aiddata_purpose%>%filter(year<=2019)%>%group_by(year)%>%summarise(total=sum(usd))%>%arrange(year)%>%ggplot()+geom_point(aes(x=year,y=total))+labs(y="Commitment Amount (USD 2011)")
```


Below we get the top 10 purpose codes and plot the amount given under each code by year, in a line graph.
```{r}
code=aiddata_purpose%>%group_by(purp_code)%>%summarise(total=sum(usd))%>%mutate(rank=min_rank(desc(total)))%>%filter(rank<=10)%>%select(purp_code)


data=aiddata_purpose%>%right_join(code,by="purp_code")%>%group_by(year,purp_code)%>%summarise(total=sum(usd))%>%filter(year<=2019)


data["purp_code"]=lapply(data["purp_code"],as.character)

data%>%ggplot+geom_line(aes(x=year,y=total,color=purp_code))+labs(y="Commitment Amount (USD 2011)", color="Purpose Code")+scale_color_brewer(palette="Paired")
```

Below we get the top 10 purpose codes and plot the amount given under each code by year, in a bar plot.
```{r}
code=aiddata_purpose%>%group_by(purp_code)%>%summarise(total=sum(usd))%>%mutate(rank=min_rank(desc(total)))%>%filter(rank<=10)%>%select(purp_code)


data=aiddata_purpose%>%right_join(code,by="purp_code")%>%group_by(year,purp_code)%>%summarise(total=sum(usd))%>%filter(year<=2019)


data["purp_code"]=lapply(data["purp_code"],as.character)

data%>%ggplot(aes(x=year,y=total,fill=purp_code))+geom_bar(stat="identity")+labs(y="Commitment Amount (USD 2011)", fill="Purpose Code")+scale_fill_brewer(palette="Spectral")


```



Below we get the top 10 purpose codes and plot the percent given of aid given under that code, by year, in a line graph.
```{r}
code=aiddata_purpose%>%group_by(purp_code)%>%summarise(total=sum(usd))%>%mutate(rank=min_rank(desc(total)))%>%filter(rank<=10)%>%select(purp_code)

year_total=aiddata_purpose%>%group_by(year)%>%summarize(year_tot=sum(usd))%>%select(year,year_tot)


data=aiddata_purpose%>%right_join(code,by="purp_code")%>%group_by(year,purp_code)%>%summarise(total=sum(usd))%>%filter(year<=2019)

data2=data%>%left_join(year_total, by="year")


data2["purp_code"]=lapply(data2["purp_code"],as.character)


data2%>%mutate(percent=total/year_tot)%>%ggplot+geom_line(aes(x=year,y=percent,color=purp_code))+labs(y="Percent of Aid", color="Purpose Code")+scale_color_brewer(palette="Paired")
```




Below we get the top 10 purpose codes and plot the percent given of aid given under that code, by year, in a stacked bar graph.


```{r}
code=aiddata_purpose%>%group_by(purp_code)%>%summarise(total=sum(usd))%>%mutate(rank=min_rank(desc(total)))%>%filter(rank<=10)%>%select(purp_code)

year_total=aiddata_purpose%>%group_by(year)%>%summarize(year_tot=sum(usd))%>%select(year,year_tot)


data=aiddata_purpose%>%right_join(code,by="purp_code")%>%group_by(year,purp_code)%>%summarise(total=sum(usd))%>%filter(year<=2019)

data2=data%>%left_join(year_total, by="year")


data2["purp_code"]=lapply(data2["purp_code"],as.character)

data3=data2%>%mutate(percent=total/year_tot)%>%select(purp_code,year,percent)


data3%>%ggplot(aes(x=year,y=percent,fill=purp_code))+geom_bar(stat="identity")+labs(y="Percent of Aid", fill="Purpose Code")+scale_fill_brewer(palette="Spectral")
```