---
title: "Mini Project 2"
author: "Nick Connolly and Jinhua Xu"
date: "5/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r libraries, echo=FALSE, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(reshape2)
```

In this project, we compare the amount of money each US state spends on health care coverage and consider the average amount spent on each person. The R code is included to help future researchers reproduce our analysis.

```{r data}
CoverageData <- read.csv(file="./1_Dataset/hcare_cov.csv")
SpendingData <- read.csv(file="./1_Dataset/hcare_spend.csv")

CombinedTable <- left_join(select(CoverageData,Location,X2013__Total,X2014__Total),select(SpendingData,Location,X2013__Total.Health.Spending,X2014__Total.Health.Spending), by = "Location" )

CombinedTableFiltered <- filter(CombinedTable, X2013__Total.Health.Spending < 100000000, X2013__Total < 100000000 )

XDataSpending <- melt( data.frame(select(CombinedTableFiltered,Location,X2013__Total,X2014__Total)), id.vars = "Location")
```

## Comparing Health Care Data by State, 2013-2014

The bar chart below shows the total amount of money spent on health care by each state in 2013 and 2014. In general, spending increased slightly in 2014.

```{r plots}

XDataSpending %>%
  ggplot( mapping = aes( x=Location, y=value, fill=variable )) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.5)) + theme(axis.text.x = element_text(angle=90,hjust=1,vjust=1)) + xlab("State") + ylab("Total Health Care Spending (USD)") + ggtitle("Health Care Spending by State, 2013-2014")

```

The average amount spent on health care per person is computed for each state. This average increases from 2013 to 2014.

```{r data coverage}

CombinedTableCoverage <- CombinedTable %>%
  mutate( Avg_Cov_2013 = X2013__Total/X2013__Total.Health.Spending ) %>%
  mutate( Avg_Cov_2014 = X2014__Total/X2014__Total.Health.Spending )

XDataCoverage <- melt( data.frame(select(CombinedTableCoverage,Location,Avg_Cov_2013,Avg_Cov_2014)), id.vars = "Location")


XDataCoverage %>%
  ggplot( mapping = aes( x=Location, y=value, fill=variable, color=variable )) +
  geom_point() + theme(axis.text.x = element_text(angle=90,hjust=1,vjust=1)) + xlab("State") + ylab("Average Coverage (USD)") + ggtitle("Comparing Health Care Coverage, 2013-2014")

```
