---
title: "Ran"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 0: Load necessary libraries and datasets
Load libraries that we use and the two healthcare datasets specified for the project.
We spent some time on the website reading about what's included in the datasets.

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(modelr)
library(broom)
library(readxl)

teen_birth <- read.csv("../Data/teen_birth.csv")

unemployment <- read_excel("../Data/Unemployment.xls", skip = 7)
```
# Step 1: Tidy data into more accessible form
Using the tidyverse functions that we've learned, clean the two datasets so that data analysis wolud be more efficient.

```{r}
teen_birth2015 <- teen_birth %>%
  filter(Year == 2015) %>%
  rename(FIPS = Combined.FIPS.Code, BirthRate = Birth.Rate) %>%
  select(State, County, FIPS, BirthRate)

unemployment2015 <- unemployment %>%
  rename(FIPS = FIPStxt) %>%
  select(FIPS, State, Area_name, 39:42) %>%
  rename(LaborForce = Civilian_labor_force_2015, Employed = Employed_2015, Unemployed = Unemployed_2015, UnemploymentRate = Unemployment_rate_2015)
  
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
