---
title: "Mini Project 1"
author: "Elisa and Shinhae"
date: "5/21/2019"
output:
  html_document: default
  pdf_document: default
---


#Introduction
For this project we used the AidData. 
Each row of this data set contains a financial transaction between two countries and from an international organization to a country. Attributes of the data set include Year, Donor, Recipient, Commitment Amount, and Coalesced Purpose Name, but for out analysis we only used Year, Donor, Recipient, Commitment Amount.

In this project we first compared the number of countries each donor donated to with the amout of money each donor donated. We were interested in finding out if countries or organizations that donate to the largest number of recipients also donate the largest amount of money. This analysys was performed over all the years present in the data set (from 1947 to 2013). It resulted that the donor which donated to more countries was United States followed by Spain, Japan, Germany; while the donor which donated the most money over the years is the International Bank for Reconstruction and Development, followed by United States, International Monetary Fund, Japan. We note that, if we ignore the donations from international organizations, the donors that donated the to more countries also donated the most amount of money.

Then we investigated how the total amount of the donations from all the donor varies over the years from 1947 to 2013. In the data set one year was recorded as "9999", so we ignored the data corresponding to such year. We discovered that the amount of the donations have been increasing over the years and reached its maximum in 2010. 

We first imported packages the AidData that we would use. 

```{r setup}
library(ggplot2)
library(tidyverse)
library(readr)
```


```{r, cache=TRUE}
setwd("/cloud/project/Day2/Mini-Project-Visualize/1_Data")
mydata<-read_csv("AidDataCoreThin_ResearchRelease_Level1_v3.1.csv")
```


# Results

#WHO DONATES? 
How are donations distributed across countries? Who donates to whom? Are there any patterns, for example some group of countries tends to donate only to some specific group of other countries? Or maybe some countries tend to receive only from a specific set of countries?
```{r, cache=TRUE}
y <- mydata %>%
  mutate(donor = fct_recode(donor,"UNICEF" = "United Nations Children`s Fund (UNICEF)" ))%>%
  select(donor, recipient)%>%
  group_by(donor)%>%
  summarise(number_recipients=n()) %>%
  arrange(desc(number_recipients)) %>%
  slice(1:9)
y
y$donor <- factor(y$donor, levels =   y$donor[order(y$number_recipients, decreasing = T)])
  #filter(number_recipients>80000)
  
y %>%
  ggplot(color = 'purple') +
  geom_bar(aes(x=donor, y= number_recipients, fill= donor), stat = 'identity')+
  theme(axis.text.x=element_text(angle=40, hjust=1, vjust=1))+ labs(x='Donors')+ labs(y= 'number of recipients')
```



```{r, cache=TRUE}
z <- mydata %>%
   mutate(donor = fct_recode(donor,"IBRD" = "World Bank - International Bank for Reconstruction and Development (IBRD)", "International Monetary Fund" =
 "International Monetary Fund (IMF)", "World Bank" = "World Bank - International Development Association (IDA)", "IADB" = "Inter-American Development Bank (IADB)"))%>%
   select(donor, commitment_amount_usd_constant)%>%
   group_by(donor)%>%
   summarise(total = sum(commitment_amount_usd_constant))%>%
   arrange(desc(total))%>%
   slice(1:9)
  #z
  
z$donor <- factor(z$donor, levels = z$donor[order(z$total, decreasing = T)])
#filter(commitment_amount_usd_constant>5000000000)
z %>%
   ggplot() +
  geom_bar(aes(x=donor,y= total, fill = donor), stat = 'identity')+
   theme(axis.text.x=element_text(angle=40, hjust=1, vjust=1))
```


#HOW MUCH DO THEY DONATE? 
How much do countries donate and receive? Who donates the most/least? Are there countries that donate and also receive? How does the amount donated/received by country change over time?

```{r, cache=TRUE}
mydata %>% 
  select(donor, recipient, commitment_amount_usd_constant) %>% 
  group_by(donor) %>% 
  summarize(max_donation=max(commitment_amount_usd_constant)) %>% 
  arrange(desc(max_donation)) %>% 
  slice(1:5)
```

Throughout the time from 1947 to 2013, the donor, International Monetary Fund (IMF), donated the most. 

<!-- ```{r} -->
<!-- mydata %>%  -->
<!--   filter(year!=9999) %>%  -->
<!--   select(year, donor, commitment_amount_usd_constant) %>%  -->
<!--   group_by(year, donor) %>%  -->
<!--   summarise(n=sum(commitment_amount_usd_constant)) %>%  -->
<!--   arrange(desc(year)) -->
<!--   slice(1:9) -->
<!-- ``` -->


```{r}
mydata %>% 
  filter(year!=9999) %>% 
  select(year, commitment_amount_usd_constant) %>% 
  group_by(year) %>% 
  summarise(n=sum(commitment_amount_usd_constant)) %>% 
  #arrange(desc(n)) %>% 
  ggplot(aes(x=year,y = n, fill = year))+
  geom_bar(stat = "identity")+
  geom_point(aes(x=2010, y=max(n)),size=0.5, color="red")+ylab("total amount")
```

The amount donated/received by country increased, in general, over time. However, we could see that it dropped drastically in 2004 and then soared in 2009 and reached the maximum in 2010. 

<!-- ```{r} -->
<!-- mydata %>%  -->
<!--   filter(year==2010) %>%  -->
<!--   select(donor, recipient, commitment_amount_usd_constant, coalesced_purpose_name) %>%  -->
<!--   arrange(desc(commitment_amount_usd_constant)) %>%  -->
<!--   slice(1:10) -->

<!-- ``` -->  

