---
title: "miniproject2"
author: "Shinhae Park"
date: "5/22/2019"
output: pdf_document
---

```{r setup}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(forcats)
```

```{r, cache=TRUE}
setwd("/cloud/project/Day3/MiniProject02-Healthcare/1_Dataset")
coverage<-read_csv("hcare_cov.csv")
spend<-read_csv("hcare_spend.csv")
```

Let us check the data and the definitions of each variable.

```{r}
head(coverage) #US total excludes Puerto Rico
```


# Medicaid: 
Includes those covered by Medicaid, Medical Assistance, Children's Health Insurance Plan (CHIP) or any kind of government-assistance plan for those with low incomes or a disability, as well as those who have both Medicaid and another type of coverage, such as dual eligibles who are also covered by Medicare.

# Medicare: 
Includes those covered by Medicare, Medicare Advantage, and those who have Medicare and another type of non-Medicaid coverage where Medicare appears to be the primary payer. Excludes seniors who also report employer-sponsored coverage and full-time work, and those covered by Medicare and Medicaid (dual eligibles).

# Employer:
Includes those covered through a current or former employer or union, either as policyholder or as dependent

# Other Public:
Includes those covered under the military or Veterans Administration

# Non-Group: 
Includes those covered by a policy purchased directly from an insurance company, etiher as policyholder or as dependent

# Uninsured:
Includes those without health insurance and those who have coverage under the Indian Health Service only.

# N/A: 
Estimates with relative standard errors greater than 30% are not provided.


```{r}
#remove N/A

coverage2=coverage[-which(coverage=="N/A",arr.ind=TRUE)[,1],]
```

```{r}
coverage2 %>% 
  select(1:7) %>% 
  filter(Location!="United States") %>%
  gather("type_2013","amount", -Location) %>% 
  group_by(type_2013) %>% 
  ggplot(aes(x=Location,y=amount))+
  geom_bar(aes(fill=type_2013),stat="identity", position = "stack")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank())+
  ylab("proportion")
```



```{r}
coverage2 %>% 
  select(1:7) %>% 
  filter(Location!="United States") %>%
  gather("type_2013","amount", -Location) %>% 
  group_by(type_2013) %>% 
  ggplot(aes(x=Location,y=amount))+
  geom_col(aes(fill=type_2013))+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank())+
  ylab("proportion")
```

Now, I want to see how the coverage changed from 2013 to 2016 in Illinois.

```{r}
cov_illinois<-coverage2 %>% 
  filter(Location=="Illinois") %>%
  gather("types","amount", -Location) %>%  
  filter(!grepl("Total",types)) %>% 
  select(-c(Location))

#sp_illinois<-spend %>% 
#  filter(Location=="Illinois")
```

```{r}
cov_illinois[,"years"]<- c(rep("2013",6),rep("2014",6),rep("2015",6),rep("2016",6))
```

```{r}
#which(grepl("Employer",cov_illinois$types))
cov_illinois$types[which(grepl("Employer",cov_illinois$types))]<-"Employer"
cov_illinois$types[which(grepl("Non-Group",cov_illinois$types))]<-"Non-Group"
cov_illinois$types[which(grepl("Medicaid",cov_illinois$types))]<-"Medicaid"
cov_illinois$types[which(grepl("Medicare",cov_illinois$types))]<-"Medicare"
cov_illinois$types[which(grepl("Other Public",cov_illinois$types))]<-"Other Public"
cov_illinois$types[which(grepl("Uninsured",cov_illinois$types))]<-"Uninsured"
```


```{r}
cov_illinois %>% 
  group_by(years) %>% 
  ggplot(aes(x=years,y=amount))+
  geom_bar(aes(fill=types),stat="identity", position = "stack")+
  theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1))+
  theme(axis.text.y=element_blank(),axis.ticks.y=element_blank())+
  ylab("proportion")+ggtitle("Illinois Coverage from 2013 to 2016")
```


# Question 1: Is there a relationship between healthcare coverage and healthcare spending in the United States?



# Question 2: How does the spending distribution change across geographic regions in the United States?



# Question 3: Does the relationship between healthcare coverage and healthcare spending in the United States change from 2013 to 2014?

```{r}
spend2013_14<-spend %>% 
  select(Location,`2013__Total Health Spending`, `2014__Total Health Spending`) %>% filter(Location=="United States")
  
```

```{r}
data2013_14<-coverage %>% 
  select(Location,`2013__Total`, `2014__Total`) %>% 
  filter(Location=="United States") %>% 
  full_join(spend2013_14, by="Location") %>% 
  ggplot(aes=("Location") )+geom_bar(stat="identity",fill="pink")
  
```